Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system with plausible mechanism

### 2024-07-19 _tex/index.tex
Built site for gh-pages

DCJ has proven effective in assessing competencies and traits
predominantly within the educational realm, as demonstrated by
\citet{Pollitt_2012b}, \citet{Jones_2015}, \citet{vanDaal_et_al_2016},
\citet{Bartholomew_et_al_2018}, \citet{Lesterhuis_2018},
\citet{Bartholomew_et_al_2020}, and \citet{Marshall_et_al_2020}.
However, its application transcends education, as exemplified by
\citet{Boonen_et_al_2020}. The methodology has also evolved to include
multiple, as opposed to pairwise comparisons
\citep{Luce_1959, Placket_1975}, and to accommodate comparisons with
ordinal outcomes \citep{Tutz_1986, Agresti_1992}. Overall, research
suggests that DCJ offers an alternative and efficient approach to
measurement and evaluation, characterized by its reliability and
validity \citep{Lesterhuis_2018, vanDaal_2020, Marshall_et_al_2020}.
Nevertheless, despite the method's widespread use, there is no clear
representation in the literature of the plausible mechanisms that
generate DCJ data. Particularly, there is no depiction of the complexity
and the underlying assumptions of the DCJ system, nor how different
assessment factors can potentially influence the observed DCJ outcome.

@@

Although several studies have examined the individual impact of these
factors on the method's reliability, including \citet{Bramley_2015},
\citet{Pollitt_2012b}, \citet{Bramley_et_al_2019},
\citet{Verhavert_et_al_2019}, \citet{Crompvoets_et_al_2022},
\citet{vanDaal_et_al_2017}, and \citet{Gijsen_et_al_2021}, to the best
of the authors' knowledge, none have provided such transparent depiction
of DCJ system and the plausible mechanisms that generate the DCJ
outcome. This study aims to fill this gap by utilizing the framework of
causal analysis and Directed Acyclic Graphs {[}DAG; \citet{Pearl_2009};
\citet{Pearl_et_al_2016}{]}. Using this framework, the study will
construct a scientific model to elucidate the underlying assumptions of
the DCJ system, providing plausible mechanisms of how the DCJ data is
generated. This model will enable researchers to draw inferences about
plausible causal relationships within the DCJ system. Furthermore, using
a minimal set of assumptions from the framework, the study will
translate the scientific model into a probabilistic statistical model,
aiming to derive statistical estimands for different targets of
inference. Ultimately, the results of this study could inform the
planning of DCJ experiments and hold significance for researchers or
analysts involved in education and assessment procedures who implement
the DCJ methodology.

\section{Theoretical framework}\label{sec-TheoreticalFramework}

\subsection{Research questions and their
estimands}\label{sec-ResearchQuestions}

\subsection{A scientific model for the DCJ procedure}\label{sec-SciMod}

\subsection{From the scientific to the Bradley-Terry-Luce
model}\label{sec-Sci2StatMod}

\section{Discussion}\label{sec-discussion}

\subsection{Limitations and further research}\label{sec-LFR}

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

>>>>

DCJ has proven effective in assessing competencies and traits
predominantly within the educational realm, as demonstrated by
\citet{Pollitt_2012b}, \citet{Jones_2015}, \citet{vanDaal_et_al_2016},
\citet{Bartholomew_et_al_2018}, \citet{Lesterhuis_2018},
\citet{Bartholomew_et_al_2020}, and \citet{Marshall_et_al_2020}.
However, its application transcends education, as exemplified by
\citet{Boonen_et_al_2020}. The methodology has also evolved to include
multiple, as opposed to pairwise comparisons
\citep{Luce_1959, Placket_1975}, and to accommodate comparisons with
ordinal outcomes \citep{Tutz_1986, Agresti_1992}. Overall, research
suggests that DCJ offers an alternative and efficient approach to
measurement and evaluation, characterized by its reliability and
validity \citep{Lesterhuis_2018, vanDaal_2020, Marshall_et_al_2020}.
Nevertheless, despite the method's widespread use, the literature does
not offer a clear representation of the plausible mechanisms that
generate the DCJ data. Particularly, there is no depiction of the
complexity and the underlying assumptions of the DCJ system, nor how
different assessment factors can potentially influence the observed DCJ
outcome.

@@

Although several studies have examined the individual impact of these
factors on the method's reliability
\citep{Bramley_2015, Pollitt_2012b, Bramley_et_al_2019, Verhavert_et_al_2019, Crompvoets_et_al_2022, vanDaal_et_al_2017, Gijsen_et_al_2021},
none, to the best of the authors' knowledge, have provided a transparent
depiction of the DCJ system and the plausible mechanisms generating the
DCJ outcome. This study aims to fill this gap by representing DCJ within
the causal analysis framework \citep{Pearl_2009, Pearl_et_al_2016}.
Using Directed Acyclic Graphs
\citep[DAG,][]{Gross_et_al_2018, Neal_2020}, the study will construct a
scientific model to elucidate the underlying assumptions of the DCJ
system, providing plausible mechanisms for how the DCJ outcome could be
generated. Next, using a minimal set of assumptions embedded in the
framework, the study will translate the scientific model into a
probabilistic statistical model. This model derives statistical
estimands to draw inferences about plausible causal relationships within
the DCJ system. Ultimately, this research aims to extend the law of
comparative judgment initially proposed by \citet{Thurstone_1927} and
provide a sound probabilistic base for the statistical analysis of DCJ
data. Consequently, this research holds significance for researchers and
analysts involved in education and assessment procedures who implement
or design DCJ experiments.

\section{Preliminaries}\label{sec-prelim}

\subsection{Causal analysis and DAGs}\label{sec-prelim-causal}

\subsection{The flow of association and
causation}\label{sec-prelim-flow}

\section{Theoretical framework}\label{sec-theory}

\subsection{A scientific model for the DCJ}\label{sec-theory-scientific}

\subsection{From the scientific to statistical
model}\label{sec-theory-statistics}

\subsection{Where does it all fit?}\label{sec-theory-where}

\section{Discussion}\label{sec-discuss}

\subsection{Findings}\label{sec-discuss-finding}

\subsection{Limitations and further
research}\label{sec-discuss-limitations}

@@

\section{Appendix}\label{sec-appendix}

\subsection{Why estimate judges appart?}\label{sec-appA}

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}
